---
title: "Chocolate Flavored Exploratory Data Analysis"
author: "Rebecca Hadi"
date: "February 3rd, 2018"
output: html_document
---

As a chocolate lover, when I saw this chocolate related data set on Kaggle, I couldn't resist the temptation to play around and see what I could learn. This notebook will walk through my typical steps of exploring a new data set.  For reference, the data I'm working with can be found at https://www.kaggle.com/rtatman/chocolate-bar-ratings/data. 


I've loaded the packages ggplot2, dplyr, and tidyr (code not shown).
```{r, message=FALSE, warning=FALSE, include = FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)

```


First, I will load the data. 

```{r, message=FALSE,warning=FALSE}
chocolate <- read.csv("/Users/bt0x/Downloads/flavors_of_cacao.csv")
```

Next, I want to look at the data itself, the structure/format, and summary values. 

```{r, message = FALSE, warning = FALSE}
head(chocolate)

str(chocolate)

summary(chocolate)

```

It's interesting that the Cocoa percent is stored as a factor opposed to a numeric value. Let's create another column that has the data stored as numeric. 


```{r, message = FALSE, warning = FALSE}
chocolate$Cocoa.Percent.Numeric <-  chocolate$Cocoa.Percent %>% 
                                    as.character(.) %>%  #convert to character b/c stored as factor
                                    gsub("%", "", .) %>% #Remove percent
                                    as.numeric(.) #Convert to numeric
                                  
```



Next, I want to look at the relationship between some of the variables.  First, I'm interested in the relationship between cocoa percentage and rating. 



### What's the relationship between rating and cocoa percent?

```{r, message = FALSE, warning = FALSE}
ggplot(data = chocolate, aes(x = Cocoa.Percent.Numeric, y = Rating)) + 
       geom_point() + 
       theme_classic() + 
       ggtitle("Rating vs. Cocoa Percent") 
```

Looking at the plot, there doesn't appear to be a strong relationship between rating and cocoa percent. We could fit a linear model to validate this.  At around 70% cocoa, which has several data points, there is a wide distribution of ratings between 1-5.  


```{r, message = FALSE, warning = FALSE}
chocolate_model <- lm(Rating ~ Cocoa.Percent.Numeric, data = chocolate)
summary(chocolate_model)

```

This isn't the most intuitive coefficient to interpret because we can't have a cocoa percent of zero (maybe that's possible somewhere, but not in this data set.) Let's try it again and re-scale the cocoa percent to centered around the mean. 


```{r, message = FALSE, warning = FALSE}
#Center and scale model by mean and standard deviation respectively
cocoa.rating.z <- (chocolate$Cocoa.Percent.Numeric - mean(chocolate$Cocoa.Percent.Numeric)) / sd(chocolate$Cocoa.Percent.Numeric) 
  
#Fit Model with centered variable
chocolate_model.z <- lm(Rating ~ cocoa.rating.z, data = chocolate)

#Show model 
summary(chocolate_model.z)

```

For the mean cocoa percentage `r round(mean(chocolate$Cocoa.Percent.Numeric),2)`$\%$, we predict the rating to be 3.19. The coefficient of the centered cocoa rating is statistically significant. For every 1 standard deviation difference in cocoa rating (which is `r sd(chocolate$Cocoa.Percent.Numeric)`), we predict the rating to decrease -0.08.  This is not a large effect size and we see that the $r^2$ is very low at 0.027, indicating that the model does not do a great job of explaining the variation in our data set. 


Now that we've seen the data, it makes more sense why the cocoa rating was scored as a factor. I'm curious to see what the most common percentage of cocoa. 


```{r, message = FALSE, warning = FALSE}
ggplot(data = chocolate, aes(x = as.factor(Cocoa.Percent.Numeric))) + 
      geom_bar() + 
      theme_classic() + 
      ggtitle("Distribution of Cocoa Percent") + 
      theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
      labs (x = "Cocoa Percent", y = "Count")

```


Wow! By far, 70% cocoa is the most common in this data set. 

There is a lot more we can dig into with this data set (e.g. do any countries produce higher rated chocolate? Is the distribution of cocoa percent similar across all countries?).  

For now, I've learned that in this data set

1. 70% cocoa is the most common chocolate. 
2. There isn't a strong relationship between cocoa percent and rating.  








